Decisions Decisions Decisions: Carrying out Latent Profile Analysis in Accordance With Best Practices Using the tidyLPA R package

### 2020-12-29 paper.tex
clean up repo

Here, we provide a walkthrough of an analysis using the tidyLPA R package.

\hypertarget{installation}{%
\subsection{Installation}\label{installation}}

@@

Here, we use a simulated dataset contains data on broad (general, and not situation- or topic-specific) interest, sense of enjoyment, and
self-efficacy from the PISA assessment.

First, let's explore the data graphically using the \{ggplot2\} package:

@@

\includegraphics{paper_files/figure-latex/unnamed-chunk-6-1.pdf}

We see that the data broadly \emph{appears} to be normally distributed, as the following statistics from the use of the psych package suggest:

@@

Entropy. \texttt{prob\_min} refers to the minimum of the diagonal of the average latent
class probabilities for most likely class membership, by assigned class. The
minimum should be as high as possible, reflecting greater classification
certainty (cases are assigned to classes they have a high probability of
belonging to; see Jung \& Wickrama, 2008); \texttt{prob\_max} refers to the maximum.
Finally. \texttt{n\_min} refers to the proportion of the sample assigned to the smallest
class (based on most likely class membership); \texttt{n\_max} to the largest. Finally,
\texttt{BLRT\_p} refers to the p-value for the bootstrapped likelihood ratio test; a
value greater than .05 indicates that the model fit better than one with one
fewer class. Additional fit statistics (other penalized likelihood values, such
as the CAIC, SABIC, and ICL), and the bootstrapped LRT test statistic, can be
obtained with the \texttt{get\_fit()} function:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{pisaUSA15 }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(broad_interest, enjoyment, self_efficacy) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{single_imputation}\NormalTok{() }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{estimate_profiles}\NormalTok{(}\DecValTok{3}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{get_fit}\NormalTok{(m1)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 1 x 18
##   Model Classes  LogLik    AIC    AWE    BIC   CAIC    CLC    KIC  SABIC     ICL
##   <dbl>   <dbl>   <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl>   <dbl>
## 1     1       3 -17399. 34826. 35080. 34919. 34933. 34799. 34843. 34874. -36442.
## # ... with 7 more variables: Entropy <dbl>, prob_min <dbl>, prob_max <dbl>,
## #   n_min <dbl>, n_max <dbl>, BLRT_val <dbl>, BLRT_p <dbl>
\end{verbatim}

@@

Last, we can create a faceted plot of two-dimensional correlation plots and unidimensional density plots for a single set of model estimates:

\includegraphics{paper_files/figure-latex/unnamed-chunk-11-1.pdf}

We can use \texttt{get\_data()} to return data in wide format when applied to an object
of class tidyProfile (one element of a tidyLPA object)

@@

>>>>

Here, we provide a walkthrough of an analysis using the tidyLPA R package.
\#\# Installation

@@

Here, we use a simulated dataset contains data on broad (general, and not situation- or topic-specific) interest, sense of enjoyment, and self-efficacy from the PISA assessment.
First, let's explore the data graphically using the \{ggplot2\} package:

@@

\includegraphics{paper_files/figure-latex/unnamed-chunk-6-1.pdf}
We see that the data broadly \emph{appears} to be normally distributed, as the following statistics from the use of the psych package suggest:

@@

Entropy. \texttt{prob\_min} refers to the minimum of the diagonal of the average latent
class probabilities for most likely class membership, by assigned class. The
minimum should be as high as possible, reflecting greater classification
certainty (cases are assigned to classes they have a high probability of
belonging to; see Jung \& Wickrama, 2008); \texttt{prob\_max} refers to the maximum.
Finally. \texttt{n\_min} refers to the proportion of the sample assigned to the smallest
class (based on most likely class membership); \texttt{n\_max} to the largest. Finally,
\texttt{BLRT\_p} refers to the p-value for the bootstrapped likelihood ratio test; a
value greater than .05 indicates that the model fit better than one with one
fewer class. Additional fit statistics (other penalized likelihood values, such
as the CAIC, SABIC, and ICL), and the bootstrapped LRT test statistic, can be
obtained with the \texttt{get\_fit()} function (processed with the tidyr R package to be in \enquote{long} rather than \enquote{wide} format):

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(tidyr)}

\NormalTok{pisaUSA15 }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(broad_interest, enjoyment, self_efficacy) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{single_imputation}\NormalTok{() }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{estimate_profiles}\NormalTok{(}\DecValTok{3}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{get_fit}\NormalTok{(m1) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{gather}\NormalTok{(stat, val)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## # A tibble: 18 x 2
##    stat              val
##    <chr>           <dbl>
##  1 Model         1      
##  2 Classes       3      
##  3 LogLik   -17399.     
##  4 AIC       34826.     
##  5 AWE       35080.     
##  6 BIC       34919.     
##  7 CAIC      34933.     
##  8 CLC       34799.     
##  9 KIC       34843.     
## 10 SABIC     34874.     
## 11 ICL      -36442.     
## 12 Entropy       0.744  
## 13 prob_min      0.750  
## 14 prob_max      0.936  
## 15 n_min         0.161  
## 16 n_max         0.669  
## 17 BLRT_val    649.     
## 18 BLRT_p        0.00990
\end{verbatim}

@@

Last, we can create a faceted plot of two-dimensional correlation plots and unidimensional density plots for a single set of model estimates:
\includegraphics{paper_files/figure-latex/unnamed-chunk-11-1.pdf}
We can use \texttt{get\_data()} to return data in wide format when applied to an object
of class tidyProfile (one element of a tidyLPA object)

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{get_data}\NormalTok{(model_}\DecValTok{4}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\NormalTok{tidyr}\OperatorTok{::}\KeywordTok{gather}\NormalTok{(Class_prob, Probability, }\KeywordTok{contains}\NormalTok{(}\StringTok{"CPROB"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}
